### 2013-05-05 FinalSpec_Annotated.tex
added specs

\documentclass[letterpaper,11pt]{article}
% Change the header if you're going to change this.

% Possible packages - uncomment to use
%\usepackage{amsmath}       % Needed for math stuff
%\usepackage{lastpage}      % Finds last page
%\usepackage{amssymb}       % Needed for some math symbols
%\usepackage{graphicx}      % Needed for graphics
\usepackage[usenames,dvipsnames]{xcolor}        % Needed for graphics and color
%\usepackage{setspace}       % Needed to set single/double spacing
\usepackage{hyperref}

%Suggested by TeXworks
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim

% Sets the page margins to 1 inch each side
\usepackage[margin=1in]{geometry}

% Uncomment this section if you wish to have a header.
\usepackage{fancyhdr} 
\pagestyle{fancy} 
\renewcommand{\headrulewidth}{0.5pt} % customise the layout... 
\lhead{CS51 Final Spec} \chead{} \rhead{April 14, 2013} 
\lfoot{} \cfoot{\thepage} \rfoot{}

\frenchspacing
\setlength\parindent{0pt}
\setlength\parskip{2ex}

\newcommand{\annot}[1]{\textbf{\color{BrickRed} [#1]}}

\begin{document}
\title{CS 51 Final Project Final Specification---{\annot{Annotated!}}}
\author{
Luis Perez $|$ luisperez@college.harvard.edu \\ 
Andy Shi $|$ andyshi@college.harvard.edu \\ 
Zihao Wang $|$ zihaowang01@college.harvard.edu \\ 
Ding Zhou $|$ dzhou@college.harvard.edu
}
\date{May 5, 2013}
\maketitle

%Put your stuff here
{\annot{To the reader: Annotations are again in this form}}.

\section{Overview}

We have decided to reduce the amount of functions available to the outside
world. The matrix module will still contain the same functionality as before,
but only the functions necessary for the implementation of the Simplex algorithm
will be exposed (row-reduction, for example). The main focus of the project has
now shifted to having a working simplex algorithm. Other functions, such as
finding the eigenvalues of a matrix, while interesting, have been placed in our
“Cool Features” category. Furthermore, following the idea that Simplex is our
main focus, additional functionality such as parsing input, providing arbitrary
float precision and/or providing a matrix implementation with bignums has been
degraded in importance. All of our attention will be focused on creating a
working simplex algorithm. {\annot{All of this happened, except for the parsing input part. We were able to parse inputs from plaintext files}}.

\section{Detailed Description}
\subsection{Signatures and Interfaces}
\subsubsection{Matrices}

Our matrix module will actually be a functor that takes in an
\verb@ORDERED_AND_OPERATIONAL@ module. The \verb@ORDERED_AND_OPERATIONAL@ module
will be defined as is below. There will be a global order for the elements, also
as defined below. In our beta implementations, we plan to simply use Ocaml
floats as the elements, but, if time allows, the plan is to make the simplex
algorithm rely on abstract as matrices a possible so that we can change the
\verb@ORDERED_AND_OPERATIONAL@ to something more efficient and/or accurate.

\begin{verbatim}
type order = Equal | Less | Greater

module type ORDERED_AND_OPERATIONAL
sig
type t
val zero : t
val one: t
val compare : t -> t -> order

(* Converts a t to a string *)
val to_string : t -> string

val add: t -> t -> t
val subtract: t -> t -> t
val multiply: t -> t -> t
val divide: t -> t -> t

(* For testing *)

(* Prints a t *)
val print: t -> unit

(* Generates the same t each time when called *)
val generate: unit -> t

(* Generates a t greater than the argument passed in *)
val generate_gt: t -> unit -> t

(* Generates a t less than the argument passed in *)
val generate_lt: t -> unit -> t

(* Generates a t in between the two arguments. Returns none if none exist *)
val generate_between: t -> t -> unit -> t option
end
\end{verbatim}

{\annot{A somewhat important change to this section was the addition of a 
generate\_x function. Our abstract elements became too abstract and we
couldn't efficiently generate numbers for testing, so we wrote this function which
basically performs a casting operation. In our actual (non-testing) code, we
wrote a from\_string function which we ended up using more}}.

The functions in the module, since they will be dedicated to matrices and
optimized for matrix elements, will not be exposed to the outside world. The
module will be passed in as a constructor to the Matrix module (defined later),
which will exposed only that while is, once again, necessary for the Simplex
module. Additionally, the \verb@ORDERED_AND_OPERATIONAL@ module will more than
likely have additional values and functions not exposed in its signature. For
example, we plan to define the “epsilon for comparison” for floats inside the
module, but we keep it hidden simply because it's not necessary anywhere else.

Users will have no direct access to this module, and therefore there are not
standards that need to be enforced (other than those in place by the interface).

The following signature is the one for the Matrix module, and it also the one
that has changed the most. Most of the functions we originally had were not
necessary for the simplex algorithm, so we have made them ``cool features''
which we will attempt to implement if we have time.

\begin{verbatim}
module type MATRIX =
sig
exception NonSquare
exception ImproperDimension

type elt
type matrix
(* represented using Ocaml's built-in Arrays *)

(* empty matrix *)
val empty

(* returns the nth row of a matrix as a matrix*)
val get_row : int -> matrix
(* implemented in terms of get_elt *)

(* returns the nth column of a matrix as a matrix *)
val get_coln : int -> matrix
(*also implemented in terms of get_elt *)

(* returns the element in the nth row and mth column *)
val get_elt : int -> int -> matrix 
(* will just get the (n,m) index of an array *)

(* Takes a list of lists and converts that to a matrix *)
val from_list : (elt list list) -> matrix
(* Will implement using nested match statements *)

(* Scales every element in the matrix by another elt *)
val scale : matrix -> elt -> matrix
(* Will implement by iterating through the matrix and scaling each element *)

(* Adds two matrices. They must have the same dimensions *)
val add : matrix -> matrix -> matrix
(* Will add the elements elementwise and construct a new matrix *)

(* Multiplies two matrices. If the matrices have dimensions m x n and p x q, n
* and p must be equal, and the resulting matrix will have dimension m x q *)
val mult: matrix -> matrix -> matrix

(* Returns the row reduced form of a matrix *)
val row_reduce: matrix -> matrix
(* We will implement the algorithm found in the link above *)

(* Prints out the contents of a matrix *)
print: matrix -> unit
(* Iterate through the matrix and print each element *)
end
\end{verbatim}

The exceptions exists in case the user tries to perform operations which require
a square matrix or matrices with certain dimensions. Matrices have their own
type, and they contain elements of the type \verb@elt@. The following is a list
of helper functions in the module, but not exposed to the outside world.

\begin{verbatim}
(* Will take the dot product of the nth row of the first matrix and the jth
 * column of the second matrix to create the n,j th entry of the resultant 
 * where the matrices are a single dimensional arrays *)
val dot -> matrix -> matrix
(* Will implement this based on the specification in the Algorithms book *)
\end{verbatim}

{\annot{We had \emph{a lot} more helper functions to get a row or column, set a
row or column, perform basic row operations, etc. They all made our lives easier
at some point in the project}}.

The following are cool features which we will try to implement if we have time.

\textbf{Cool Features}:

\begin{verbatim}
(* Returns the inverse of a matrix *)
val inverse: matrix -> matrix

(* Returns the norm of the matrix *)
val norm: matrix -> elt

(* Transposes a matrix. If the input has dimensions m x n, the output will
 * have dimensions n x m *)
val transpose: matrix -> matrix
(* Will basically ``flip'' the indices of the input matrix

(* Returns the trace of the matrix *)
val trace: matrix -> elt
(* Will check if the matrix is square, then sum up all the elements along its
 * diagonal *)

(* Returns the determinant of the matrix *)
val det: matrix -> elt
(* Will implement this algorithm based on a description in Hubbard. Involves
 * column reducing the input (or row-reducing the transpose) and then keeping
 * track of the operations to build a sequence of coefficients to multiply *)
{\annot{We were able to implement this in time, using LU decomposition}}

(* Returns a list of eigenvalues and eigenvectors of a matrix *)
val eigen: matrix -> (elt *matrix) list option
(* Calculates successive powers of the input matrix, each multiplied by the
* same basis vector. Generates a polynomial and solves for zeros, which
* yields eigenvalues. Repeat for all basis vectors *) 
\end{verbatim}

{\annot{We were able to implement inverse, transpose, trace, and determinant.
The first 3 functions we were able to implement as specified, but the
determinant was implemented by decomposing it into upper and lower triangular
matrices}}

\subsubsection{Simplex Algorithm}

The simplex solves linear programs, which are basically linear inequalities.
Given inequalities, using the simplex algorithm, we can find the maximum value
of a given expression. What the simplex algorithm does is that it takes in
several linear inequalities and writes them so that they are equalities. For
instance $5x + 7y \leq 10$ would be $5x + 7y + s = 10$ where $s \geq 0$ is the
slack variable. Then it takes all these equations and puts them a a matrix where
each entry represents the coefficient of the variable in the expression. The
first row represents the equation we are optimizing. The final column is the
constant values of each equation. Then we will do a series of computations so
that the top row has all nonnegative values. This can be done using the
functions from the Matrix module.

The steps are as follows:

Look along the first row and find the smallest negative element. Look at the
column of that entry and compute the value/element of each row. If any element
is zero, we can ignore it. Otherwise take the element with the smallest
value/element and make this the pivot point. Reduce the column so that the pivot
point in 1 and any other entry in the column is zero. Repeat the process again
until the top row consists of only nonnegative values. The first entry in the
last column is the maximum. In addition, for any column consisting of all zeroes
and a one, we can find the value of the corresponding variable by looking at the
value of the row containing the one. If the column doesn’t consists of zeroes
and a one, then the value of the corresponding variable is zero.

This algorithm also works with variables that are negative, inequalities where
the unknowns is greater than a constant, and minimum of the equation. This can
be easily done by either multiplying the whole inequality by -1 or by negating
the variable. If calculating the minimum or negating the variable, we can revert
them later.

\begin{verbatim}
module type simplex
sig
type elt
type linear_equation
type constraint

val pivot : matrix -> matrix

(* Will implement these if we have time *)
val constraint_from_string : string -> constraint
val linear_from_string : string -> constraint
end
\end{verbatim}

These are helper functions that we think will be used:

\verb@no_neg : matrix -> bool@

Check the top row has no negative elements, which can be done by using the next
helper function

\verb@min_of_row : row -> elt*(elt ref)@

A helper for the row reduction which will find the minimum element and its
location.

{\annot{We again had a ton more helper functions, plus we did not anticipate
some edge cases in the simplex algorithm. One such case was the fact that our
algorithm relied on guessing an initial solution to the optimization problem
which works for a majority of cases. However, if the constraint equations are
set up correctly, our initial solution would fail to satisfy the constraint
equations. We then had to follow a complex procedure outlined in the
\emph{Algorithms} book to resolve this}}.

\subsection{Modules and Actual Code}
Please see our attached files or our repo at 
\url{www.github.com/Fantastic-Four/ocaml-matrix}. Files of note include Elts.ml
and Matrix.ml.

\subsection{Timeline}
\begin{itemize}
\item Week 1 (4/7--4/13):
  \begin{itemize}
    \item Signature definitions 
    \item Setting up Github (everyone on the same page)
    \item Writing makefile
    \item Establishing work distribution
  \end{itemize}

\item Week 2 (4/14--4/20):
  \begin{itemize}
    \item Basic matrix functionality
    \item Some rudimentary form of matrix multiplication and row reduction
    \item Finishing up the element types
  \end{itemize}

\item Week 3 (4/21--4/27):
  \begin{itemize}
    \item Beta testing the simplex algorithm
    \item Implementation of simple input parser for testing (we will most likely
be inputting arrays)
    \item Completion of the matrix module algorithms
  \end{itemize}

\item Week 4 (4/28--5/4)
  \begin{itemize}
    \item Fully working simplex algorithm
    \item Implement a better matrix multiplication algorithm
    \item Add some of the cool features!
  \end{itemize}
\end{itemize}

\subsection{Progress Report}
We have already implemented the design structure for all of the files
(specifically, each file has a signature detailing what it needs to do). The
Makefile is also ready, while rudimentary functionality has been implemented in
the matrix module allowing most of the functions we want matrices to have to be
available already. For evidence of our progress please look at Elts.ml and
Matrix.ml.

\section{Version Control}
We are currently using github for version control. Our repository can be found
here, under the organization Fantastic Four
(\url{www.github.com/Fantastic-Four/ocaml-matrix}).We had some difficulties at
first getting the repo set up on everyone’s computers but we fixed that over the
weekend and we should be good to go.

\end{document}

### 2013-05-05 DraftSpec_Annotated.tex

\documentclass[letterpaper,11pt]{article}
% Change the header if you're going to change this.

% Possible packages - uncomment to use
%\usepackage{amsmath}       % Needed for math stuff
%\usepackage{lastpage}      % Finds last page
%\usepackage{amssymb}       % Needed for some math symbols
%\usepackage{graphicx}      % Needed for graphics
\usepackage[usenames,dvipsnames]{xcolor}        % Needed for graphics and color
%\usepackage{setspace}       % Needed to set single/double spacing
\usepackage{hyperref}

%Suggested by TeXworks
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim

% Sets the page margins to 1 inch each side
\usepackage[margin=1in]{geometry}

% Uncomment this section if you wish to have a header.
\usepackage{fancyhdr} 
\pagestyle{fancy} 
\renewcommand{\headrulewidth}{0.5pt} % customise the layout... 
\lhead{CS51 Draft Spec} \chead{} \rhead{April 7, 2013} 
\lfoot{} \cfoot{\thepage} \rfoot{}

\frenchspacing
\setlength\parindent{0pt}
\setlength\parskip{2ex}

\newcommand{\annot}[1]{\textbf{\color{BrickRed} [#1]}}

\begin{document}
\title{CS 51 Final Project Draft Specification---\annot{Annotated!}}
\author{
Luis Perez $|$ luisperez@college.harvard.edu \\ 
Andy Shi $|$ andyshi@college.harvard.edu \\ 
Zihao Wang $|$ zihaowang01@college.harvard.edu \\ 
Ding Zhou $|$ dzhou@college.harvard.edu
}
\date{May 5, 2013}
\maketitle

%Put your stuff here
{\annot{Note for the reader: This is the annotated version of our CS 51 
final project specification. All annotations will appear in this color. Our changes to the actual modules and the finer points of implementation were too numerous to annotate (so there are no annotations in section 3)}}.

\section{Brief Overview}

Our project seeks to implement an efficient matrix library for Ocaml. The end
goal is to utilize this matrix library in conjunction with the simplex algorithm
to find solution to problems in linear programming. In order to accomplish this,
there will be multiple small checkpoints. We need to implement and efficient
matrix representation (so as to not use the naive implementation of lists of
lists). We also hope to provide a useful interface that allows the user to input
arbitrarily large matrices effectively (possible by mapping over the entries of
the matrix). One big challenge for the matrix library will be the implementation
of efficient algorithms for multiplication and row reduction. We expect the
multiplication to initially consists of the standard algorithm, but will
eventually grow into either block multiplication or Strassen {\annot{We were not
able to implement more sophisticated multiplication algorithms}}. We might even
implement some way to chose between the algorithms depending on the inputted
matrices, but that would not an essential feature {\annot{We were not able to do
this either}}.

The row reduction algorithm will certainly be one of the most fundamental
algorithms in our matrix library implementation. From our current understanding
of simplex, it seems like it will heavily rely on row reduction {\annot{The
simplex algorithm ended up relying on a different type of row reduction, so we
couldn't just plug in a matrix into our row reduction algorithm. We had to do a
lot of extra work to get simplex}}. The only algorithm we can come up with so
far is Gauss-Jordan, though we have found some algorithms which are more suited
for computation by computers with imprecise storage of floats {\annot{In fact,
we just had to be careful which element to choose as a pivot}}.

The simplex part of the problem is the main goal, but if we can get down the
matrix module, then it appears as if the simplex algorithm should be
straightforward {\annot{This was definitely not the case}}. Eventually, we wish
to implement a simplex module that allows the input of arbitrary linear
equations with linear constraints, and then we parse this information into the
corresponding matrix and finally perform the simplex algorithm on that matrix in
order to return a solution to the specified problem.

Other tentative goals are to implement matrices that can accept either
arbitrarily large numbers (bignums) or arbitrarily precise floats 
(\verb@int list * int stream@). It would also be interesting, though not
necessary, to implement multiple algorithms for each operation and choose from
them the one that is the best for a specific input data {\annot{We were not able
to implement any of this functionality}}.

\section{Feature List}
\begin{enumerate}

\item Find a way to represent a matrix (arrays or list list or other
representation)---will likely use an array representation using the Ocaml array
module (\url{http://caml.inria.fr/pub/docs/manual-ocaml/libref/Array.html})
{\annot{Implemented as specified}}

\item Make a matrix from a list of lists (this is how the user will input the
data of the matrix). {\annot{Implemented as specified}}

\item Scalar multiplication---as described here
(\url{http://www.purplemath.com/modules/mtrxmult.htm})
{\annot{Implemented as specified}}

\item Matrix addition---as described here
(\url{http://www.purplemath.com/modules/mtrxmult.htm})
{\annot{Implemented as specified}}

\item Standard Matrix multiplication
(\url{http://www.purplemath.com/modules/mtrxmult.htm}) 
{\annot{Implemented as specified}}

\item Row Reduction (Gauss-Jordan). We will use the following resource to help
us calculate row reduction when the entries of the matrix are stored with
limited precision
(\url{http://thejuniverse.org/PUBLIC/LinearAlgebra/LOLA/rowRed/var.html})
{\annot{Implemented as specified}}

\item Matrix inverse / Row reduction with LU Decomposition - Description can be
found here (\url{http://www.math.ust.hk/~macheng/math111/LU_Decomposition.pdf})
{\annot{To find the inverse, we augmented the original matrix with an identity
matrix, then row-reduced the augmented matrix}}

\item Simplex Algorithm (Described in Algorithms book, in addition to the
resources provided below)
{\annot{Implemented mostly as specified, with some modifications from the
Wikipedia article on simplex}}

\item Matrix norm 
{\annot{Did not implement}}

\item Matrix transpose
{\annot{Implemented as specified}}

\item Trace---This is a typical trace (sum of the entries on the diagonal)
{\annot{Implemented as specified}}

\item Determinant---found using algorithm described in Section 4.8 of Hubbard's
\emph{Vector Calculus, Linear Algebra, and Differential Forms: A Unified
Approach} (Math 23a/b textbook)
{\annot{Implemented by decomposing the matrix into upper and lower triangular
matrices, then multiplying the entries along the diagonals for both and
multiplying the two results}}

\item Eigenvectors / Eigenvalues, calculated also by a method in Hubbard's book
(this might not be necessary to carry out the simplex algorithm and we might
abandon it if we can't figure out how to implement it in a timely manner) 
{\annot{Did not implement}}

\end{enumerate}

\section{Draft Technical Specification}

\subsection{Matrices}

Our matrix module will actually be a functor that takes in an
\verb@ORDERED_AND_OPERATIONAL@ module (from ps4 and moogle) so we can have
matrices of ints, floats, or floats as streams.

The following type definition (a la ps4) will give us a way to talk about order
for different data types.

\begin{verbatim}
type order = Equal | Less | Greater
\end{verbatim}

The following signature will allow us to provide matrix functionality on a
variety of data types. For example, we could make an
\verb@ORDERED_AND_OPERATIONAL@ type for floats, where add would be defined as
(+.), for example. In our matrix functor, we will then use the functions defined
in the signature below (for example, add instead of (+.)). Also, we have
included some generating functions to help us with testing.

\begin{verbatim}
module type ORDERED_AND_OPERATIONAL
sig
  type t

  val zero : t
  val one: t
  val compare : t -> t -> order

  (* Converts a t to a string *)
  val to_string : t -> string

  val add: t -> t -> t
  val subtract: t -> t -> t
  val multiply: t -> t -> t
  val divide: t -> t -> t

  (* For testing *)
  (* Prints a t *)
  val print: t -> unit

  (* Generates the same t each time when called *)
  val generate: unit -> t

  (* Generates a t greater than the argument passed in *)
  val generate_gt: t -> unit -> t

  (* Generates a t less than the argument passed in *)
  val generate_lt: t -> unit -> t

  (* Generates a t in between the two arguments. Returns none if none exist *)
  val generate_between: t -> t -> unit -> t option
end
\end{verbatim}

The following signature will provide all the necessary functions that can be
performed on a matrix. We include basic matrix operations like additions,
multiplications, scalar multiplication, and then standard operations like
finding the inverse of a matrix, finding its trace, determinant, eigenvalues,
transpose, norm, and how to row reduce the matrix.

Below each function, in comments, is a brief description of how we will
implement the function.

\begin{verbatim}
module type MATRIX = 
sig
  exception NonSquare

  type elt
  type matrix
  (* Type of this is unknown, but will probably be represented using Ocaml’s
   * built-in Arrays *)
  (* empty matrix *)
  val empty

  (* Takes a list of lists and converts that to a matrix *)
  val from_list : (elt list list) -> matrix
  (* Will implement using nested match statements *)

  (* Scales every element in the matrix by another elt *)
  val scale : matrix -> elt -> matrix
  (* Will implement by iterating through the matrix and scaling each element *)

  (* Adds two matrices. They must have the same dimensions *)
  val add : matrix -> matrix -> matrix
  (* Will add the elements elementwise and construct a new matrix *)

  (* Multiplies two matrices. If the matrices have dimensions m x n and p x q, n
   * and p must be equal, and the resulting matrix will have dimension m x q *)
  val mult: matrix -> matrix -> matrix
  (* Will take the dot product of the nth row of the first matrix and the jth
   * column of the second matrix to create the n,j th entry of the resultant *)

  (* Returns the row reduced form of a matrix *)
  val row_reduce: matrix -> matrix 
  (* We will implement the algorithm found in the link above *)

  (* Returns the inverse of a matrix *)
  val inverse: matrix -> matrix
  (* Will implement this based on the specification in the Algorithms book *)

  (* Returns the norm of the matrix *)
  val norm: matrix -> elt

  (* Transposes a matrix. If the input has dimensions m x n, the output will
   * have dimensions n x m *)
  val transpose: matrix -> matrix
  (* Will basically ``flip’’ the indices of the input matrix

  (* Returns the trace of the matrix *)
  val trace: matrix -> elt
  (* Will check if the matrix is square, then sum up all the elements along its
   * diagonal *)

  (* Returns the determinant of the matrix *)
  val det: matrix -> elt
  (* Will implement this algorithm based on a description in Hubbard. Involves
   * column reducing the input (or row-reducing the transpose) and then keeping 
   * track of the operations to build a sequence of coefficients to multiply *)

  (* Returns a list of eigenvalues and eigenvectors of a matrix *)
  val eigen: matrix -> (elt *matrix) list option
  (* Calculates successive powers of the input matrix, each multiplied by the
   * same basis vector. Generates a polynomial and solves for zeros, which 
   * yields eigenvalues. Repeat for all basis vectors *)

  (* Takes a string and builds a matrix from it *)
  from_string : string -> matrix
  (* We will have some way to express matrices using strings, and then we will
   * parse the string to give the matrix *)

  (* Prints out the contents of a matrix *)   
  print :matrix -> unit
  (* Iterate through the matrix and print each element *)
end
\end{verbatim}

The exception exists in case the user tries to perform operations, such as trace
or determinant, which require a square matrix. Matrices have their own type, and
they contain elements of the type \verb@elt@.

We will probably need a helper function to raise matrices to powers and to solve
for zeros of a polynomial if we want to implement the \verb@eigen@ function.

\subsection{Simplex Algorithm}

The simplex solves linear programs, which are basically linear inequalities.
Given inequalities, using the simplex algorithm, we can find the maximum value
of a given expression. What the simplex algorithm does is that it takes in
several linear inequalities and writes them so that they are equalities. For
instance $5x + 7y \leq 10$ would be $5x + 7y + s = 10$ where $s \geq 0$ is the
slack variable. Then it takes all these equations and puts them a a matrix where
each entry represents the coefficient of the variable in the expression. The
first row represents the equation we are optimizing. The final column is the
constant values of each equation. Then we will do a series of computations so
that the top row has all nonnegative values. This can be done using the
functions from the Matrix module.

The steps are as follows:

Look along the first row and find the smallest negative element. Look at the
column of that entry and compute the value/element of each row. If any element
is zero, we can ignore it. Otherwise take the element with the smallest
value/element and make this the pivot point. Reduce the column so that the pivot
point in 1 and any other entry in the column is zero. Repeat the process again
until the top row consists of only nonnegative values. The first entry in the
last column is the maximum. In addition, for any column consisting of all zeroes
and a one, we can find the value of the corresponding variable by looking at the
value of the row containing the one. If the column doesn’t consists of zeroes
and a one, then the value of the corresponding variable is zero.

This algorithm also works with variables that are negative, inequalities where
the unknowns is greater than a constant, and minimum of the equation. This can
be easily done by either multiplying the whole inequality by -1 or by negating
the variable. If calculating the minimum or negating the variable, we can revert
them later.

\begin{verbatim}
module type simplex
sig
  type elt
  type linear_equation
  type constraint

  val pivot : matrix -> matrix
  val constraint_from_string : string -> constraint
  val linear_from_string : string -> constraint
end
\end{verbatim}

These are helper functions that we think will be used:

\verb@no_neg : matrix -> bool@

Check the top row has no negative elements, which can be done by using the next
helper function

\verb@min_of_row : row -> elt*(elt ref)@

A helper for the row reduction which will find the minimum element and its
location.

\section{What's Next}

We will read more into the algorithms and get a basic understanding of how they
all work so we can better judge the difficulty of each algorithm. We will also
read more into OCaml to see how to put together a project with multiple files
(since we have always relied on staff’s frameworks for each pset). We will
working through the CS50 appliance. We already have setup a git repository on
Github (\url{https://github.com/Fantastic-four}), and most of us are fairly
familiar with version control.

\end{document}

### 2013-05-05 report.tex
Addinging info.txt, readme, and report

--- 2013-05-05 DraftSpec_Annotated.tex
@@
 %\usepackage{amsmath}       % Needed for math stuff
-%\usepackage{lastpage}      % Finds last page
+%\usepackage{lastpage}      % Finds last 
 %\usepackage{amssymb}       % Needed for some math symbols
@@
 \begin{document}
-\title{CS 51 Final Project Draft Specification---\annot{Annotated!}}
+\title{CS 51 Final Project Report---\annot{Annotated!}}
 \author{
@@
 
-%Put your stuff here
-{\annot{Note for the reader: This is the annotated version of our CS 51 
-final project specification.
-All annotations will appear in this color.
-Our changes to the actual modules and the finer points of implementation were too numerous to annotate (so there are no annotations in section 3)}}.
-
 \section{Brief Overview}
-
-Our project seeks to implement an efficient matrix library for Ocaml. The end
-goal is to utilize this matrix library in conjunction with the simplex algorithm
-to find solution to problems in linear programming. In order to accomplish this,
-there will be multiple small checkpoints. We need to implement and efficient
-matrix representation (so as to not use the naive implementation of lists of
-lists). We also hope to provide a useful interface that allows the user to input
-arbitrarily large matrices effectively (possible by mapping over the entries of
-the matrix). One big challenge for the matrix library will be the implementation
-of efficient algorithms for multiplication and row reduction. We expect the
-multiplication to initially consists of the standard algorithm, but will
-eventually grow into either block multiplication or Strassen {\annot{We were not
-able to implement more sophisticated multiplication algorithms}}. We might even
-implement some way to chose between the algorithms depending on the inputted
-matrices, but that would not an essential feature {\annot{We were not able to do
-this either}}.
-
-The row reduction algorithm will certainly be one of the most fundamental
-algorithms in our matrix library implementation. From our current understanding
-of simplex, it seems like it will heavily rely on row reduction {\annot{The
-simplex algorithm ended up relying on a different type of row reduction, so we
-couldn't just plug in a matrix into our row reduction algorithm. We had to do a
-lot of extra work to get simplex}}. The only algorithm we can come up with so
-far is Gauss-Jordan, though we have found some algorithms which are more suited
-for computation by computers with imprecise storage of floats {\annot{In fact,
-we just had to be careful which element to choose as a pivot}}.
-
-The simplex part of the problem is the main goal, but if we can get down the
-matrix module, then it appears as if the simplex algorithm should be
-straightforward {\annot{This was definitely not the case}}. Eventually, we wish
-to implement a simplex module that allows the input of arbitrary linear
-equations with linear constraints, and then we parse this information into the
-corresponding matrix and finally perform the simplex algorithm on that matrix in
-order to return a solution to the specified problem.
-
-Other tentative goals are to implement matrices that can accept either
-arbitrarily large numbers (bignums) or arbitrarily precise floats 
-(\verb@int list * int stream@). It would also be interesting, though not
-necessary, to implement multiple algorithms for each operation and choose from
-them the one that is the best for a specific input data {\annot{We were not able
-to implement any of this functionality}}.
 
@@
 module (\url{http://caml.inria.fr/pub/docs/manual-ocaml/libref/Array.html})
-{\annot{Implemented as specified}}
-
-\item Make a matrix from a list of lists (this is how the user will input the
-data of the matrix). {\annot{Implemented as specified}}
-
-\item Scalar multiplication---as described here
-(\url{http://www.purplemath.com/modules/mtrxmult.htm})
-{\annot{Implemented as specified}}
-
-\item Matrix addition---as described here
-(\url{http://www.purplemath.com/modules/mtrxmult.htm})
-{\annot{Implemented as specified}}
-
-\item Standard Matrix multiplication
-(\url{http://www.purplemath.com/modules/mtrxmult.htm}) 
-{\annot{Implemented as specified}}
-
-\item Row Reduction (Gauss-Jordan). We will use the following resource to help
-us calculate row reduction when the entries of the matrix are stored with
-limited precision
-(\url{http://thejuniverse.org/PUBLIC/LinearAlgebra/LOLA/rowRed/var.html})
-{\annot{Implemented as specified}}
-
-\item Matrix inverse / Row reduction with LU Decomposition - Description can be
-found here (\url{http://www.math.ust.hk/~macheng/math111/LU_Decomposition.pdf})
-{\annot{To find the inverse, we augmented the original matrix with an identity
-matrix, then row-reduced the augmented matrix}}
-
-\item Simplex Algorithm (Described in Algorithms book, in addition to the
-resources provided below)
-{\annot{Implemented mostly as specified, with some modifications from the
-Wikipedia article on simplex}}
-
-\item Matrix norm 
-{\annot{Did not implement}}
-
-\item Matrix transpose
-{\annot{Implemented as specified}}
-
-\item Trace---This is a typical trace (sum of the entries on the diagonal)
-{\annot{Implemented as specified}}
-
-\item Determinant---found using algorithm described in Section 4.8 of Hubbard's
-\emph{Vector Calculus, Linear Algebra, and Differential Forms: A Unified
-Approach} (Math 23a/b textbook)
-{\annot{Implemented by decomposing the matrix into upper and lower triangular
-matrices, then multiplying the entries along the diagonals for both and
-multiplying the two results}}
-
-\item Eigenvectors / Eigenvalues, calculated also by a method in Hubbard's book
-(this might not be necessary to carry out the simplex algorithm and we might
-abandon it if we can't figure out how to implement it in a timely manner) 
-{\annot{Did not implement}}
 
@@
 
-Our matrix module will actually be a functor that takes in an
-\verb@ORDERED_AND_OPERATIONAL@ module (from ps4 and moogle) so we can have
-matrices of ints, floats, or floats as streams.
-
-The following type definition (a la ps4) will give us a way to talk about order
-for different data types.
-
-\begin{verbatim}
-type order = Equal | Less | Greater
-\end{verbatim}
-
-The following signature will allow us to provide matrix functionality on a
-variety of data types. For example, we could make an
-\verb@ORDERED_AND_OPERATIONAL@ type for floats, where add would be defined as
-(+.), for example. In our matrix functor, we will then use the functions defined
-in the signature below (for example, add instead of (+.)). Also, we have
-included some generating functions to help us with testing.
-
-\begin{verbatim}
-module type ORDERED_AND_OPERATIONAL
-sig
-  type t
-
-  val zero : t
-  val one: t
-  val compare : t -> t -> order
-
-  (* Converts a t to a string *)
-  val to_string : t -> string
-
-  val add: t -> t -> t
-  val subtract: t -> t -> t
-  val multiply: t -> t -> t
-  val divide: t -> t -> t
-
-  (* For testing *)
-  (* Prints a t *)
-  val print: t -> unit
-
-  (* Generates the same t each time when called *)
-  val generate: unit -> t
-
-  (* Generates a t greater than the argument passed in *)
-  val generate_gt: t -> unit -> t
-
-  (* Generates a t less than the argument passed in *)
-  val generate_lt: t -> unit -> t
-
-  (* Generates a t in between the two arguments. Returns none if none exist *)
-  val generate_between: t -> t -> unit -> t option
-end
-\end{verbatim}
-
-The following signature will provide all the necessary functions that can be
-performed on a matrix. We include basic matrix operations like additions,
-multiplications, scalar multiplication, and then standard operations like
-finding the inverse of a matrix, finding its trace, determinant, eigenvalues,
-transpose, norm, and how to row reduce the matrix.
-
-Below each function, in comments, is a brief description of how we will
-implement the function.
-
-\begin{verbatim}
-module type MATRIX = 
-sig
-  exception NonSquare
-
-  type elt
-  type matrix
-  (* Type of this is unknown, but will probably be represented using Ocaml’s
-   * built-in Arrays *)
-  (* empty matrix *)
-  val empty
-
-  (* Takes a list of lists and converts that to a matrix *)
-  val from_list : (elt list list) -> matrix
-  (* Will implement using nested match statements *)
-
-  (* Scales every element in the matrix by another elt *)
-  val scale : matrix -> elt -> matrix
-  (* Will implement by iterating through the matrix and scaling each element *)
-
-  (* Adds two matrices. They must have the same dimensions *)
-  val add : matrix -> matrix -> matrix
-  (* Will add the elements elementwise and construct a new matrix *)
-
-  (* Multiplies two matrices. If the matrices have dimensions m x n and p x q, n
-   * and p must be equal, and the resulting matrix will have dimension m x q *)
-  val mult: matrix -> matrix -> matrix
-  (* Will take the dot product of the nth row of the first matrix and the jth
-   * column of the second matrix to create the n,j th entry of the resultant *)
-
-  (* Returns the row reduced form of a matrix *)
-  val row_reduce: matrix -> matrix 
-  (* We will implement the algorithm found in the link above *)
-
-  (* Returns the inverse of a matrix *)
-  val inverse: matrix -> matrix
-  (* Will implement this based on the specification in the Algorithms book *)
-
-  (* Returns the norm of the matrix *)
-  val norm: matrix -> elt
-
-  (* Transposes a matrix. If the input has dimensions m x n, the output will
-   * have dimensions n x m *)
-  val transpose: matrix -> matrix
-  (* Will basically ``flip’’ the indices of the input matrix
-
-  (* Returns the trace of the matrix *)
-  val trace: matrix -> elt
-  (* Will check if the matrix is square, then sum up all the elements along its
-   * diagonal *)
-
-  (* Returns the determinant of the matrix *)
-  val det: matrix -> elt
-  (* Will implement this algorithm based on a description in Hubbard. Involves
-   * column reducing the input (or row-reducing the transpose) and then keeping 
-   * track of the operations to build a sequence of coefficients to multiply *)
-
-  (* Returns a list of eigenvalues and eigenvectors of a matrix *)
-  val eigen: matrix -> (elt *matrix) list option
-  (* Calculates successive powers of the input matrix, each multiplied by the
-   * same basis vector. Generates a polynomial and solves for zeros, which 
-   * yields eigenvalues. Repeat for all basis vectors *)
-
-  (* Takes a string and builds a matrix from it *)
-  from_string : string -> matrix
-  (* We will have some way to express matrices using strings, and then we will
-   * parse the string to give the matrix *)
-
-  (* Prints out the contents of a matrix *)   
-  print :matrix -> unit
-  (* Iterate through the matrix and print each element *)
-end
-\end{verbatim}
-
-The exception exists in case the user tries to perform operations, such as trace
-or determinant, which require a square matrix. Matrices have their own type, and
-they contain elements of the type \verb@elt@.
-
-We will probably need a helper function to raise matrices to powers and to solve
-for zeros of a polynomial if we want to implement the \verb@eigen@ function.
-
-\subsection{Simplex Algorithm}
-
-The simplex solves linear programs, which are basically linear inequalities.
-Given inequalities, using the simplex algorithm, we can find the maximum value
-of a given expression. What the simplex algorithm does is that it takes in
-several linear inequalities and writes them so that they are equalities. For
-instance $5x + 7y \leq 10$ would be $5x + 7y + s = 10$ where $s \geq 0$ is the
-slack variable. Then it takes all these equations and puts them a a matrix where
-each entry represents the coefficient of the variable in the expression. The
-first row represents the equation we are optimizing. The final column is the
-constant values of each equation. Then we will do a series of computations so
-that the top row has all nonnegative values. This can be done using the
-functions from the Matrix module.
-
-The steps are as follows:
-
-Look along the first row and find the smallest negative element. Look at the
-column of that entry and compute the value/element of each row. If any element
-is zero, we can ignore it. Otherwise take the element with the smallest
-value/element and make this the pivot point. Reduce the column so that the pivot
-point in 1 and any other entry in the column is zero. Repeat the process again
-until the top row consists of only nonnegative values. The first entry in the
-last column is the maximum. In addition, for any column consisting of all zeroes
-and a one, we can find the value of the corresponding variable by looking at the
-value of the row containing the one. If the column doesn’t consists of zeroes
-and a one, then the value of the corresponding variable is zero.
-
-This algorithm also works with variables that are negative, inequalities where
-the unknowns is greater than a constant, and minimum of the equation. This can
-be easily done by either multiplying the whole inequality by -1 or by negating
-the variable. If calculating the minimum or negating the variable, we can revert
-them later.
-
-\begin{verbatim}
-module type simplex
-sig
-  type elt
-  type linear_equation
-  type constraint
-
-  val pivot : matrix -> matrix
-  val constraint_from_string : string -> constraint
-  val linear_from_string : string -> constraint
-end
-\end{verbatim}
-
-These are helper functions that we think will be used:
-
-\verb@no_neg : matrix -> bool@
-
-Check the top row has no negative elements, which can be done by using the next
-helper function
-
-\verb@min_of_row : row -> elt*(elt ref)@
-
-A helper for the row reduction which will find the minimum element and its
-location.
-
-\section{What's Next}
-
-We will read more into the algorithms and get a basic understanding of how they
-all work so we can better judge the difficulty of each algorithm. We will also
-read more into OCaml to see how to put together a project with multiple files
-(since we have always relied on staff’s frameworks for each pset). We will
-working through the CS50 appliance. We already have setup a git repository on
-Github (\url{https://github.com/Fantastic-four}), and most of us are fairly
-familiar with version control.
-
 \end{document}

### 2013-05-05 DraftSpec_Annotated.tex
minor changes to spec
minor changes in draft spec, added newlines to print statements in Elt

@@
 %Put your stuff here
-{\annot{Note for the reader: This is the annotated version of our CS 51 
-final project specification.
-All annotations will appear in this color.
-Our changes to the actual modules and the finer points of implementation were too numerous to annotate (so there are no annotations in section 3)}}.
+{\annot{Note for the reader: This is the annotated version of our CS 51 final
+project specification. All annotations will appear in this color. Our changes to
+the actual modules and the finer points of implementation were too numerous to
+annotate (so there are no annotations in section 3)}}.
 
@@
   val transpose: matrix -> matrix
-  (* Will basically ``flip’’ the indices of the input matrix
+  (* Will basically ``flip'' the indices of the input matrix
 
@@
 and a one, we can find the value of the corresponding variable by looking at the
-value of the row containing the one. If the column doesn’t consists of zeroes
+value of the row containing the one. If the column doesn't consists of zeroes
 and a one, then the value of the corresponding variable is zero.

### 2013-05-05 DraftSpec_Annotated.tex
Finished annotating Draft ie functional specs

--- 2013-05-05 DraftSpec_Annotated.tex
@@
 \renewcommand{\headrulewidth}{0.5pt} % customise the layout... 
-\lhead{CS51 Draft Spec} \chead{} \rhead{April 7, 2013} 
+\lhead{CS51 Draft Spec} \chead{} \rhead{May 5th, 2013} 
 \lfoot{} \cfoot{\thepage} \rfoot{}
@@
 \begin{document}
-\title{CS 51 Final Project Draft Specification---\annot{Annotated!}}
+\title{CS 51 Final Project Functional Specification{\annot{:Matrix Module and Simplex Algorithm}}
+---{\annot{Annotated!}}}
 \author{
@@
 {\annot{Note for the reader: This is the annotated version of our CS 51 
-final project specification.
-All annotations will appear in this color.
-Our changes to the actual modules and the finer points of implementation were too numerous to annotate (so there are no annotations in section 3)}}.
+Final Project Functional Specifications. All annotations will appear in this color.}}.
 
@@
 goal is to utilize this matrix library in conjunction with the simplex algorithm
-to find solution to problems in linear programming. In order to accomplish this,
+to find solution {\annot{Linear Programs}}. In order to accomplish this,
 there will be multiple small checkpoints. We need to implement and efficient
 matrix representation (so as to not use the naive implementation of lists of
-lists). We also hope to provide a useful interface that allows the user to input
+lists){\annot{(Successfully accomplished. Our implementation uses array for efficienty)}}.
+ We also hope to provide a useful interface that allows the user to input
 arbitrarily large matrices effectively (possible by mapping over the entries of
-the matrix). One big challenge for the matrix library will be the implementation
+the matrix){\annot{The user can input matrices with a simple text file composed of comma
+seperated values}}. One big challenge for the matrix library will be the implementation
 of efficient algorithms for multiplication and row reduction. We expect the
@@
 eventually grow into either block multiplication or Strassen {\annot{We were not
-able to implement more sophisticated multiplication algorithms}}. We might even
-implement some way to chose between the algorithms depending on the inputted
-matrices, but that would not an essential feature {\annot{We were not able to do
-this either}}.
+able to implement more sophisticated multiplication algorithms, but it seems that the current
+algorithm works fairly well}}. We might even implement some way to chose between the 
+algorithms depending on the inputted matrices, but that would not an essential feature 
+{\annot{This feature can still easily be implemented, but there simply was no time. We
+already knew this, though, by the time we wrote our Technical Specs}}.
 
@@
 couldn't just plug in a matrix into our row reduction algorithm. We had to do a
-lot of extra work to get simplex}}. The only algorithm we can come up with so
+lot of extra work to get simplex. Though, many of the helper functions for row
+reduction ended up being extremely useful for Simplex, so this was non as pointless
+as some might thing}}. The only algorithm we can come up with so
 far is Gauss-Jordan, though we have found some algorithms which are more suited
 for computation by computers with imprecise storage of floats {\annot{In fact,
-we just had to be careful which element to choose as a pivot}}.
+we just had to be careful which element to choose as a pivot. Futhermore, we
+ended up implementing arbitrary precision, so this became a non-issue}}.
 
@@
 corresponding matrix and finally perform the simplex algorithm on that matrix in
-order to return a solution to the specified problem.
+order to return a solution to the specified problem.{\annot{This is fully functionalt.
+Input can be made in the form of a text file with a linear program. The algorithm
+runs and not only returns the right values for simple cases, but also checks for
+cases that might initially appear to be unsolvable but aren't, and for cases which
+are unbounded.}}
 
@@
 necessary, to implement multiple algorithms for each operation and choose from
-them the one that is the best for a specific input data {\annot{We were not able
-to implement any of this functionality}}.
+them the one that is the best for a specific input data {\annot{We were able to
+implement arbitrary precisions and arbitrarily large sizes with the OCaml Nums
+library}}.
 
@@
 
+\item{\annot{Make a matrix from a text file. The user is now able to read in a
+text file and create a matrix.}}
+
+\item{\annot{For Information of further functionality, take a look at MatrixI.ml.}}
+
 \item Simplex Algorithm (Described in Algorithms book, in addition to the
@@
 \item Matrix norm 
-{\annot{Did not implement}}
+{\annot{Did not implement since it was not necessary for Simplex.}}
 
 \item Matrix transpose
-{\annot{Implemented as specified}}
+{\annot{Implemented as specified.}}
 
 \item Trace---This is a typical trace (sum of the entries on the diagonal)
-{\annot{Implemented as specified}}
+{\annot{Implemented as specified.}}
 
@@
 abandon it if we can't figure out how to implement it in a timely manner) 
-{\annot{Did not implement}}
+{\annot{Did not implement since it was not necssary for Simplex.}}
 
@@
 matrices of ints, floats, or floats as streams. 
+{\annot{This actually ended up being very helpful since at the end we we able 
+to use OCaml's Num Library in order to achieve arbitrary precision and size.}}
 
@@
 
+{\annot{For the updated view of this, please take a look at EltsI.ml}}
 \begin{verbatim}
@@
 
+{\annot{For the updated view of this, please take a look at MatrixI.ml}}
 \begin{verbatim}
@@
 
+{\annot{For the updated view of this, please take a look at SimplexI.ml}}
 \begin{verbatim}
@@
 location.
+{\annot{We ended up needing many more helper functions than those shown here.}}
 

### 2013-05-05 FinalSpec_Annotated.tex
Finished annotating Draft ie functional specs

@@
 \renewcommand{\headrulewidth}{0.5pt} % customise the layout... 
-\lhead{CS51 Final Spec} \chead{} \rhead{April 14, 2013} 
+\lhead{CS51 Final Spec} \chead{} \rhead{May 5th, 2013} 
 \lfoot{} \cfoot{\thepage} \rfoot{}
@@
 \begin{document}
-\title{CS 51 Final Project Final Specification---{\annot{Annotated!}}}
+\title{CS 51 Final Project Technical Specification: Matrix Module and Simplex Algorithm---{\annot{Annotated!}}}
 \author{
@@
 %Put your stuff here
-{\annot{To the reader: Annotations are again in this form}}.
+{\annot{Note for the reader: This is the annotated version of our CS 51 
+technical project specification. All annotations will appear in this color. }}.
 
@@
 but only the functions necessary for the implementation of the Simplex algorithm
-will be exposed (row-reduction, for example). The main focus of the project has
+will be exposed (row-reduction, for example{\annot{Row Reduction ended up not
+being necessary. We have instead exposed a large set of library functions and,
+additionally, some functions specific to Simplex that would not be exposed otherwise.
+All of this is documented in MatrixI.ml}}). The main focus of the project has
 now shifted to having a working simplex algorithm. Other functions, such as

### 2013-05-05 FinalSpec_Annotated.tex

@@
 degraded in importance. All of our attention will be focused on creating a
-working simplex algorithm. {\annot{All of this happened, except for the parsing input part. We were able to parse inputs from plaintext files}}.
+working simplex algorithm. {\annot{All of this happened, even the parsing input part. 
+We were able to parse inputs from plaintext files}}.
 
@@
 \verb@ORDERED_AND_OPERATIONAL@ to something more efficient and/or accurate.
+{\annot{In the end, we ended up using the Nums OCaml library in order to obtain
+arbitrary precison and size with our elements. We made this decision late in the
+coding, but because we coded everything abstractly, changing from Floats to Nums
+was as simple as exposing the right files and setting the right modules}}
 
+{\annot{Updated version of this can be found in EltsI.ml}}
 \begin{verbatim}
@@
 module, but we keep it hidden simply because it's not necessary anywhere else.
-
-Users will have no direct access to this module, and therefore there are not
-standards that need to be enforced (other than those in place by the interface).
+{\annot{This was mostly implemented as described.}}
 
@@
 
+{\annot{This Module changed significantly. For an updated look, see MatrixI.ml}}
 \begin{verbatim}
@@
 determinant was implemented by decomposing it into upper and lower triangular
-matrices}}
+matrices. The functions that weren't implemented were mostly left behind because
+they were not essential to the Simplex Algorithm}}
 

### 2013-05-05 report.tex
adding stuff to specs

@@
 \renewcommand{\headrulewidth}{0.5pt} % customise the layout... 
-\lhead{CS51 Draft Spec} \chead{} \rhead{April 7, 2013} 
+\lhead{CS51 Final Proect Report} \chead{} \rhead{May 5, 2013} 
 \lfoot{} \cfoot{\thepage} \rfoot{}
@@
 \begin{document}
-\title{CS 51 Final Project Report---\annot{Annotated!}}
+\title{CS 51 Final Project Report}
 \author{
@@
 
-\section{Brief Overview}
+\section{Overview}
 
-\section{Feature List}
-\begin{enumerate}
+In the real world, people love to make money. Even in virtual worlds (bitcoin?),
+people love to make money. Basically, many problems in business and even
+everyday life (but especially business), boil down to maximizing profit or
+minimizing costs. Of course, when you speak of minimization or maximization
+there are always constraints. Resources are finite and time is precious.
+Therefore, in an abstract sense, the problem boils down to minimizing or
+maximizing some quantity under certain constraints. For example, you might want
+to maximize the number of textbooks x,y,z you buy under the constraints that you
+can't buy more than five from one company, the total cost can't be more
+than \$500, and the last two books you buy must cost more than \$75. This problem
+is precisely the type of problems that our program can now solve. How did we do
+this?
 
-\item Find a way to represent a matrix (arrays or list list or other
-representation)---will likely use an array representation using the Ocaml array
-module (\url{http://caml.inria.fr/pub/docs/manual-ocaml/libref/Array.html})
+To start off, we created a framework. We implemented a matrix module with
+standard matrix operations such as multiplication, addition, map (where scaling
+follows) as well as many extra features such as row reduction, determinant,
+trace, and many more. The matrix module basically allows a user to do anything
+imaginable with a matrix, in addition to providing read and write functionality
+for easy loading of data, and printing functionality for straightforward viewing
+of the abstract matrix. However the intellectually challenging part of our
+project is the implementation of the simplex algorithm. We needed to create a
+matrix module simply because OCaml doesn't provide one. We use many of the
+functions in the matrix module to create and manipulate simplex systems within
+our simplex module. This allowed us to abstract away yet another part of our
+project.
 
-\end{enumerate}
+The simplex algorithm is a type of linear programming that allows one to
+optimize a linear equation given linear constraints (like the example problem
+at the beginning). In mathematical terms, the simplex program operates on
+linear programs in standard form. That is, it seeks to minimize (or maximize)
+c.x subject to $A.x = b$ (Wikipedia). Let's consider the inequalities: $30x +
+15y + 18z \geq 500, x \geq 5,y \geq 5, z \geq 5$, and $15y+18z \geq 75$ and the objective
+functions $x +y + z$ where $x, y, z \geq 0$. As you will notice, this is precisely
+the system of equations necessary to solve our book example above. The
+algorithm converts the inequalities into equalities by adding positive slack
+variables. Then it represents this set of equations in the matrix form where
+the coefficients of the terms we maximize are on the first row and each column
+represents a variable. This provides affordable storage of most of the
+information necessary to solve the linear program. Furthermore, we then define
+basic and non-basic variables (kept track of in a list), which together with
+the aforementioned matrix, create a simplex system.
 
-\section{Draft Technical Specification}
+The first function necessary for simplex is called \verb@simple_solve@. This
+function, in abstract terms, finds the minimum value of a passed in system (as
+described above). When we call the function \verb@simple_solve@, the program
+finds the entering and leaving variables which are the variable that would
+exchange position from basic and non-basic (slack to non-slack in the initial
+case). Here, the algorithm is making the assumption that the system has a
+feasible solution where all of our non-basic variables are zero, and therefore
+exchanges the role of one of these variables with the corresponding basic
+variable. In geometric terms, we are moving along one edge of the n-gon formed
+by the constraints and attempting to minimize. This step is programmatically
+accomplished by the helper function pivot. The function \verb@simple_solve@
+continues to loop (moving along edges and minimizing) until the first
+row---which represents the term we want to optimize consists of only nonpositive
+values or there is no possible entering variable (ie, any change in the
+variables will only increase our answer, not minimize it). In the first case,
+algebraically, it means that the constraint can only increase if the variables
+change. Therefore, we have found our minimum value which is the constant stored
+at the end of the first row. This is the optimized value of the objective
+function. In the second case, where no suitable entering variable can be found,
+it means that the system can be decreased infinitely as pleased and therefore it
+is Unbounded. In the case where a solution exists, one possible value for the
+original variables can also be found by looking at their corresponding columns
+and checking to see if they are basic. If there are an infinite number of these,
+our algorithm returns only one.
 
-\subsection{Matrices}
+What if we had the constraint $x + y + z \geq -10$. As stated above, our
+\verb@simple_solve@ makes the assumption that one of the feasible solutions is
+the zero solution, but if we initially set $x, y$, and $z$ to zero the
+constraint does not hold. We therefore have a helper function called
+\verb@initialize_simplex@ that converts everything into the ``simple'' form so
+we can just recursively call \verb@simple_solve@. It does this by following a
+very complex algorithm better detailed below. Our final  simplex algorithm now
+accounts for all cases as long as the equations are linear. It also handles not
+only less than or equal constraints, but also greater than or equal to and
+equality constraints.  Finally, not only can it minimize, but it can also
+maximize. You can follow the algorithm at each step on www.simplexme.com, but we
+found that SimplexMe only works for the ``simple'' case. It has errors for large
+matrices and the infeasible and unbounded cases.
+
+\section{Planning}
+
+We planned to have a completed Simplex solver for simple cases, handling less
+than or equal constraints in addition to a complete Matrix module. Here are our
+initial functional spec and technical spec. We achieved far more than we
+expected to achieve.
+
+Our matrix module is fairly robust and it certainly implements everything that
+is necessary to program the simplex algorithm (which was our main goal for this
+module). Yet, it goes further. We managed to implement determinant, trace,
+row-reduction,inverse, LU decomposition, and more (all extra functions that we
+classified under our ``cool features'' category. There were only minor
+points that we did not achieve, such as the eigen function.
+
+As to the Elts module, we really went far and beyond. Because we coded
+everything from an abstract point of view, once we had completed our project we
+were able to go back and add a new Nums module which provided arbitrary
+precision and size (ie, implementing arbitrary float arithmetic and bignum
+arithmetic). Of course, to be honest, we really didn't actually code most of
+this up. We used the built in Nums library. But as can be seen inside of
+Elts.ml, our code was planned out so well that we can replaced the foundation of
+mathematics (\verb@ORDERED_AND_OPERATIONAL@) with only a few changes in code.
+
+The Simplex module also went further than expected (though it certainly was more
+difficult  than expected too). Not only did we implement the ability to solve
+simple linear programs, but we can now solve with certainty any linear program
+(dual and primal), and provide the use with not only the optimal value, but also
+one possible solution (which, in most cases is the only solution).
+
+Most of the milestones were achieved for all of our individual modules, with the
+exception of our \verb@row_reduce@ (which was buggy and took longer than
+expected), and of our simplex module (we went a day over what we had originally
+planned).
+
+\section{Design and Implementation}
+
+Implementing simplex was difficult. A lot of difficulty stemmed from the
+complexity and size of the algorithm itself, and we spent a lot of time trying
+to understand each step. We never completely internalized every step of the
+algorithm which made debugging difficult because it took a lot of effort to
+recall the correct steps. We also had some initial terminology confused which
+caused some of the code to be written incorrectly. Since each step ended up
+being so complex, whenever we moved to the following steps of the algorithm we
+always had trouble recalling what exactly we were doing and if someone else had
+already done it. Furthermore, bugs were very difficult to find because the
+system was so complex and there weren't many cases for which we could work
+out the steps by hand.
+
+Finding SimplexMe was essential to finally getting all of the bugs out of our
+system, but it also added many hours of unnecessary debugging. SimplexMe has
+implemented simplex incorrectly and therefore the answer it gives for non-simple
+systems can be, at times, incorrect. We verified this with Mathematica's
+built in Linear Programming function.
+
+Testing the system also proved to be a problem in its own right. Even after we
+got all of the parsing functions down and were certain that the data was being
+loaded correctly, we were never really able to find a large set of simplex
+example problems. In the end, we resorted to each team member individually
+finding a problem and solution (with Mathematica, of course, since SimplexMe
+proved unreliable), and then typing out the data into a text file so we could
+test it. Once we had all the data, we decided to program an interface that would
+allow for expanding the testability of the code. More information on this can be
+found the the \verb@README@.
+
+As to the design and implementation of our actual code base, it definitely can
+be improved. It is the best it can be given the fact that we hadn't already done
+this same problem before, but as with anything, it can be better. For example,
+our \verb@initialize_simplex@ function ended up being very complex and
+complicated. We all understand the general idea of what it does, but not all of
+us could implement it by ourselves from scratch. Basically,
+\verb@initialize_simplex@ takes in a system and adds the slack variables to it.
+During this step, if the system appears unfeasible, it adds an extra variable
+and attempts to minimize that (ie, replaces the objective function). If this new
+system has a minimum of zero, then the original system is feasible and all we
+need to do is substitute the old objective back in appropriately. Otherwise, the
+original system is unfeasible and the program returns None. In the first case,
+an extra pivot operation might be necessary, but again, initialize simplex takes
+care of this.  In the end, it calls \verb@simple_solve@ (the function we
+originally wrote) on this now simple case and a solution is produced. However,
+we did run our simplex algorithm against an equation with 100 variables and 100
+constraints (\verb@test22.txt@), and it solved it in under 1 second. Even though
+the worst case running time of simplex is not polynomial (it's exponential), on
+most real world cases simplex runs very quickly, as evidenced by our test.
+
+Each member of the team contributed to the project. The work was split between
+the matrix and elts modules, parsing the data, and the simplex algorithm. Andy
+and Luis implemented most of the matrix module and parsing function, along with
+the general framework of the program that allowed us to efficiently run and test
+our code. Luis, Jason, and Ding worked on implementing and debugging the simplex
+algorithm. Ding focused on comprehending every step of the algorithm as
+presented in the Algorithms book and then presenting that information in a more
+comprehensive manner to the the group, while Jason focused on understanding the
+process behind simplex and on coming up with example problems by hand. Jason and
+Ding worked on testing simplex, and Luis and Andy provided much of the
+functionality that allowed for easy testing.
+
+\section{Reflection}
+
+Our original planning went pretty well and we accomplished most of what we
+expected to accomplish. We mostly hit both our milestones, though we missed our
+second milestone by a day. We were definitely surprised by how difficult simplex
+ended up being, and there were times when we thought we were essentially done
+with the algorithm and it turned out that we were only halfway there.
+
+Given more time, we would like to polish up input and outputs of our algorithm,
+including a more sophisticated parsing function and a user friendly output.
+Within the matrix module itself, there are a few functions that we would have
+liked to implement including finding the eigenvectors/eigenvalues and
+calculating the norm of a matrix, among others.
+
+We made the right choice when we decided to drop some of our unnecessary matrix
+module functions to spend more time on simplex, but we made an early mistake of
+not bothering to understand simplex until the latter half of the project, and
+rather than understanding the whole algorithm before we started coding, we chose
+to only understand the steps we were directly implementing, leading to the
+restructuring of major portions of our code towards the end. The next time we
+tackle an algorithm-centric project like this, we should definitely plan out the
+entire algorithm before we code (including all the possible edge cases that
+might pop up).
+
+The most important thing we learned is how important good style, readability,
+and careful planning is to programming and debugging. There were many times when
+we were confused by what we wrote when we went back to debug our code, and since
+no one was able to fully internalize every part of the algorithm, it was
+important to abstract over helper functions and different parts of the code.
+Having had someone who understood the algorithm entirely (or who had maybe
+implemented something on this scale before), would have been extremely helpful
+as it would have allowed us to better plan our time and resources as well as
+abstract away many of the more detailed concepts.
+
+\section{Advice for Future Students}
+
+Start as early as possible and put some serious work in. Don't leave the video
+until the last minute. This is a big project and showing it off is a very
+rewarding part of it, especially when done well. Also try to understand the
+general outline for every step before serious implementation, even if it means
+delaying the implementation. Designing the framework is crucial and it is easier
+to get it right the first time. Make sure to write test functions as you write
+functions and to test each function right after you finish writing it. This way,
+you will avoid having to hunt down a small bug in one obscure part of your code.
+You'll know exactly when and where your code went wrong.
+
+Additionally, make sure your group members don't play too much League of Legends
+or Starcraft!
+
+But lastly---enjoy the project and be proud of what you've accomplished!
 

### 2013-05-06 DraftSpec_Annotated.tex
adding stuff to specs
fixed specs

@@
 \begin{document}
-\title{CS 51 Final Project Functional Specification{\annot{:Matrix Module and Simplex Algorithm}}
+\title{CS 51 Final Project Functional Specification{\annot{Matrix Module and Simplex Algorithm}}
 ---{\annot{Annotated!}}}
@@
 lot of extra work to get simplex. Though, many of the helper functions for row
-reduction ended up being extremely useful for Simplex, so this was non as pointless
-as some might thing}}. The only algorithm we can come up with so
-far is Gauss-Jordan, though we have found some algorithms which are more suited
-for computation by computers with imprecise storage of floats {\annot{In fact,
-we just had to be careful which element to choose as a pivot. Futhermore, we
-ended up implementing arbitrary precision, so this became a non-issue}}.
+reduction ended up being extremely useful for Simplex, so this was not as
+pointless as some might think}}. The only algorithm we can come up with so far
+is Gauss-Jordan, though we have found some algorithms which are more suited for
+computation by computers with imprecise storage of floats {\annot{In fact, we
+just had to be careful which element to choose as a pivot. Futhermore, we ended
+up implementing arbitrary precision, via Ocaml's num library,  so this became a
+non-issue}}.
 

### 2013-05-06 report.tex
fixed specs

@@
-\documentclass[letterpaper,11pt]{article}
+\documentclass[letterpaper,12pt]{article}
 % Change the header if you're going to change this.
@@
 
+\newcommand{\vect}[1]{\mathbf{#1}}
+
 \begin{document}
@@
 
+Link to demo video: \url{}
+
 \section{Overview}
@@
 maximizing some quantity under certain constraints. For example, you might want
-to maximize the number of textbooks x,y,z you buy under the constraints that you
+to maximize the number of textbooks $x,y,z$ you buy under the constraints that you
 can't buy more than five from one company, the total cost can't be more
 than \$500, and the last two books you buy must cost more than \$75. This problem
-is precisely the type of problems that our program can now solve. How did we do
+is precisely the type of problem that our program can now solve. How did we do
 this?
@@
 follows) as well as many extra features such as row reduction, determinant,
-trace, and many more. The matrix module basically allows a user to do anything
-imaginable with a matrix, in addition to providing read and write functionality
-for easy loading of data, and printing functionality for straightforward viewing
-of the abstract matrix. However the intellectually challenging part of our
-project is the implementation of the simplex algorithm. We needed to create a
-matrix module simply because OCaml doesn't provide one. We use many of the
-functions in the matrix module to create and manipulate simplex systems within
-our simplex module. This allowed us to abstract away yet another part of our
-project.
+trace, and many more. The matrix module basically allows a user to do almost
+anything imaginable with a matrix. In addition, we provided read and write
+functionality for easy loading of data, and printing functionality for
+straightforward viewing of the abstract matrix. However the intellectually
+challenging part of our project is the implementation of the simplex algorithm.
+We needed to create a matrix module simply because OCaml doesn't provide one
+(also because we just wanted to implement a matrix module, of course). We use
+many of the functions in the matrix module to create and manipulate simplex
+systems within our simplex module. This allowed us to abstract away yet another
+part of our project.
 
 The simplex algorithm is a type of linear programming that allows one to
-optimize a linear equation given linear constraints (like the example problem
-at the beginning). In mathematical terms, the simplex program operates on
-linear programs in standard form. That is, it seeks to minimize (or maximize)
-c.x subject to $A.x = b$ (Wikipedia). Let's consider the inequalities: $30x +
-15y + 18z \geq 500, x \geq 5,y \geq 5, z \geq 5$, and $15y+18z \geq 75$ and the objective
-functions $x +y + z$ where $x, y, z \geq 0$. As you will notice, this is precisely
-the system of equations necessary to solve our book example above. The
+optimize a linear equation given linear constraints (like the example problem at
+the beginning). In mathematical terms, the simplex program operates on linear
+programs in standard form. That is, it seeks to minimize (or maximize)
+$C\vect{x}$ subject to $A\vect{x} = \vect{b} $ (Wikipedia). $\vect{x}$ and
+$\vect{b}$ are vectors. $x$ contains the variables in the equation we want to
+optimize (the objective), and $b$ contains the constants in the constraints. $C$
+and $A$ are matrices, and they represent the coefficients in the objective and
+the constraints, respectively. Let's consider the inequalities: $30x + 15y + 18z
+\geq 500, x \geq 5,y \geq 5, z \geq 5$, and $15y+18z \geq 75$ and the objective
+functions $x +y + z$ where $x, y, z \geq 0$. As you will notice, this is
+precisely the system of equations necessary to solve our book example above. The
 algorithm converts the inequalities into equalities by adding positive slack
-variables. Then it represents this set of equations in the matrix form where
-the coefficients of the terms we maximize are on the first row and each column
+variables. Then it represents this set of equations in the matrix form where the
+coefficients of the terms we maximize are on the first row and each column
 represents a variable. This provides affordable storage of most of the
 information necessary to solve the linear program.  Furthermore, we then define
-basic and non-basic variables (kept track of in a list), which together with
-the aforementioned matrix, create a simplex system.
+basic and non-basic variables (kept track of in a list), which together with the
+aforementioned matrix, create a simplex system.
 
@@
 variable. In geometric terms, we are moving along one edge of the n-gon formed
-by the constraints and attempting to minimize. This step is programmatically
-accomplished by the helper function pivot. The function \verb@simple_solve@
-continues to loop (moving along edges and minimizing) until the first
-row---which represents the term we want to optimize consists of only nonpositive
-values or there is no possible entering variable (ie, any change in the
-variables will only increase our answer, not minimize it). In the first case,
-algebraically, it means that the constraint can only increase if the variables
-change. Therefore, we have found our minimum value which is the constant stored
-at the end of the first row. This is the optimized value of the objective
-function. In the second case, where no suitable entering variable can be found,
-it means that the system can be decreased infinitely as pleased and therefore it
-is Unbounded. In the case where a solution exists, one possible value for the
-original variables can also be found by looking at their corresponding columns
-and checking to see if they are basic. If there are an infinite number of these,
-our algorithm returns only one.
+by the constraints and attempting to minimize. This step is accomplished by the
+helper function pivot. The function \verb@simple_solve@ continues to loop
+(moving along edges and minimizing) until the first row---which represents the
+term we want to optimize consists of only non-positive values or there is no
+possible entering variable (ie, any change in the variables will only increase
+our answer, not minimize it). In the first case, algebraically, it means that
+the constraint can only increase if the variables change. Therefore, we have
+found our minimum value which is the constant stored at the end of the first
+row. This is the optimized value of the objective function. In the second case,
+where no suitable entering variable can be found, it means that the system can
+be decreased infinitely as pleased and therefore it is Unbounded. In the case
+where a solution exists, one possible value for the original variables can also
+be found by looking at their corresponding columns and checking to see if they
+are basic. If there are an infinite number of these, our algorithm returns only
+one.
 
-What if we had the constraint $x + y + z \geq -10$. As stated above, our
+What if we had the constraint $x + y + z \geq -10$? As stated above, our
 \verb@simple_solve@ makes the assumption that one of the feasible solutions is
-the zero solution, but if we initially set $x, y$, and $z$ to zero the
-constraint does not hold. We therefore have a helper function called
-\verb@initialize_simplex@ that converts everything into the ``simple'' form so
-we can just recursively call \verb@simple_solve@. It does this by following a
-very complex algorithm better detailed below. Our final  simplex algorithm now
-accounts for all cases as long as the equations are linear. It also handles not
-only less than or equal constraints, but also greater than or equal to and
-equality constraints.  Finally, not only can it minimize, but it can also
-maximize. You can follow the algorithm at each step on www.simplexme.com, but we
-found that SimplexMe only works for the ``simple'' case. It has errors for large
-matrices and the infeasible and unbounded cases.
+the zero solution, but if we initially set $x, y$, and $z$ to 0 (our algorithm
+generates a ``basic solution'' by doing this step) the constraint does not hold.
+We therefore have a helper function called \verb@initialize_simplex@ that
+converts everything into the ``simple'' form so we can just recursively call
+\verb@simple_solve@. It does this by following a very complex algorithm better
+detailed below. Our final  simplex algorithm now accounts for all cases as long
+as the equations are linear. It also handles not only less than or equal
+constraints, but also greater than or equal to and equality constraints.
+Finally, not only can it minimize, but it can also maximize. You can follow the
+algorithm at each step on www.simplexme.com (we used this to debug parts of our
+code), but we found that SimplexMe only works for the ``simple'' case. It has
+errors for large matrices and the infeasible and unbounded cases.
 
 \section{Planning}
+Here is our \href{./DraftSpec_Annotated.pdf}{functional spec} and our
+\href{./FinalSpec_Annotated.pdf}{technical spec} (the links open new PDFs).
 
@@
 row-reduction,inverse, LU decomposition, and more (all extra functions that we
-classified under our ``cool features'' category. There were only minor
-points that we did not achieve, such as the eigen function.
+classified under our ``cool features'' category. There were only minor points
+that we did not achieve, such as the eigen function, which finds eigenvectors
+and eigenvalues.
 
@@
 this up. We used the built in Nums library. But as can be seen inside of
-Elts.ml, our code was planned out so well that we can replaced the foundation of
-mathematics (\verb@ORDERED_AND_OPERATIONAL@) with only a few changes in code.
+Elts.ml, our code was planned out so well that we can replace the foundation of
+the mathematics in our program (\verb@ORDERED_AND_OPERATIONAL@) with only a few
+changes in code.
 
 The Simplex module also went further than expected (though it certainly was more
-difficult  than expected too). Not only did we implement the ability to solve
-simple linear programs, but we can now solve with certainty any linear program
-(dual and primal), and provide the use with not only the optimal value, but also
-one possible solution (which, in most cases is the only solution).
+difficult to implement than we expected). Not only did we implement the ability
+to solve simple linear programs, but we can now solve with certainty any linear
+program (dual and primal---dual requires \verb@initialize_simplex@ while primal
+is the ``simple'' case), and provide the user with not only the optimal value,
+but also one possible solution (which, in most cases is the only solution).
 
@@
 allow for expanding the testability of the code. More information on this can be
-found the the \verb@README@.
+found in the \verb@README@.
+
+To use the simplex algorithm, we first parse a standardized plain text file
+containing the parameters of the objective and the constraints. The first line
+of this file must either be \verb@max@ or \verb@min@, signalling a maximization
+or minimization step. The next line denotes the objective function. The
+coefficients of each of the variables in the objective are separated by columns
+(decimals must be represented by fractions because we are using Ocaml's num
+library). The last element in this is the ``constant'' (5 if the objective is
+$2x + 3y + 5$). The next line must read \verb@subject to@ to help us with
+parsing. The next lines are the constraints, where the coefficient of each
+variable that appears in the objective must appear (in the same order as in the
+objective), separated by commas. We also allow the user to enter if the
+constraint is a $\geq, \leq$, or $=$ condition. Again, the last element is the
+constant (5 in the case of $x + y \leq 5$). It is assumed that all the variables
+in the objective function are $\geq 0$.
+
+Our program will parse this text file and generate a matrix representing the
+objective and the constraints. Additionally, our program adds slack variables
+(for example, (we change the inequalities in the constraints to equations---$3x
++ 2y \leq 5$ becomes $3x + 2y + s = 5$, where $s \geq 0$). This makes the
+problem easier to solve).
+
+The simplex algorithm basically takes the objective function and attempts to
+find a ``basic solution'' which will satisfy all the constraints. Even if this
+solution is trivial (like setting all the variables in the objective to 0), we
+take this solution and keep optimizing it by replacing a variable not in the
+objective function with a variable that is. We check the objective to make sure
+none of the coefficients are negative; if they are, we are finished and we have
+the optimum value in the top right corner of the matrix (because we can set all
+the variables in the objective to 0 and not violate the constraints). Our
+``simple'' case is the case where we want to minimize the objective and all the
+constraints are $\leq$ relations. However, we can change any problem to this
+form. Maximization can be achieved by minimizing the negative of the objective.
+Similarly, a $\geq$ constraint can be changed to a $\leq$ constraint by
+multiplying each side of the inequality by $-1$. Finally, an $=$ constraint can
+be changed into a $\geq$ and a $\leq$.
 
@@
 \verb@initialize_simplex@ takes in a system and adds the slack variables to it.  
-During this step, if the system appears unfeasible, it adds an extra variable
-and attempts to minimize that (ie, replaces the objective function). If this new
-system has a minimum of zero, then the original system is feasible and all we
-need to do is substitute the old objective back in appropriately. Otherwise, the
-original system is unfeasible and the program returns None. In the first case,
-an extra pivot operation might be necessary, but again, initialize simplex takes
-care of this.  In the end, it calls \verb@simple_solve@ (the function we
-originally wrote) on this now simple case and a solution is produced. However,
-we did run our simplex algorithm against an equation with 100 variables and 100
-constraints (\verb@test22.txt@), and it solved it in under 1 second. Even though
-the worst case running time of simplex is not polynomial (it's exponential), on
-most real world cases simplex runs very quickly, as evidenced by our test.
+During this step, if the algorithm determines a system (the objective
+function and the constraints) to be unfeasible, it adds an extra variable and
+attempts to minimize that (ie, replaces the objective function). If this new
+system has a minimum of 0, then the original system is feasible and all we need
+to do is substitute the old objective back in appropriately. Otherwise, the
+original system is unfeasible and the program returns \verb@None@. In the first
+case, an extra pivot operation might be necessary, but again,
+\verb@initialize_simplex@ takes care of this. In the end, it calls
+\verb@simple_solve@ (the function we originally wrote) on this now simple case
+and a solution is produced. However, we did run our simplex algorithm against an
+equation with 100 variables and 100 constraints (\verb@test22.txt@), and it
+solved it in under 1 second. Even though the worst case running time of simplex
+is not polynomial (it's exponential), on most real world cases simplex runs very
+quickly, as evidenced by our test.
 
@@
 presented in the Algorithms book and then presenting that information in a more
-comprehensive manner to the the group, while Jason focused on understanding the
+comprehensive manner to the group, while Jason focused on understanding the
 process behind simplex and on coming up with example problems by hand. Jason and
@@
 general outline for every step before serious implementation, even if it means
-delaying the implementation. Designing the framework is crucial and it is easier
-to get it right the first time. Make sure to write test functions as you write
-functions and to test each function right after you finish writing it. This way,
-you will avoid having to hunt down a small bug in one obscure part of your code.
-You'll know exactly when and where your code went wrong.
+delaying the implementation. Designing the framework is crucial and it makes the
+whole process easier to get right the first time. Make sure to write test
+functions as you write functions and to test each function right after you
+finish writing it. This way, you will avoid having to hunt down a small bug in
+one obscure part of your code.  You'll know exactly when and where your code
+went wrong.
 
